Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
revise

\newcommand{\judgeCreduce}[2]{#1 \longmapsto_{\mathcal{x}} #2}
\newcommand{\judgeCreduceTrans}[2]{#1 \longmapsto_{\mathcal{x}}^{*} #2}
\newcommand{\judgeCeval}[2]{eval_{\mathcal{x}}(#1) = #2}
\newcommand{\redrule}[3]{#1 & \longmapsto_\mathcal{x} & #2 & #3\\}
\newcommand{\hiredrule}[3]{\highlight{#1} & \highlight{\longmapsto_\mathcal{x}} 

@@

\newcommand{\ineffCEK}{$\mathcal{C}$}

@@

We start by defining a CEK machine \effCEK{C} that is parameterized
over the cast representation $C$
(Section~\ref{sec:framework:cek}). This machine is space-efficient
provided that the cast representation performs compression.  The
interface between the machine and the cast representation is defined
by an abstract data type, named Cast ADT, that we define shortly 
(Section~\ref{sec:framework:adt}).  We
conjecture that all cast representations defined in the literature
are instances of this ADT.

@@

@@

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
  \label{def:cast-rep}
  A cast abstract data type is a set $Cast$, which is indexed by two types, 
  with four operators:
  \begin{description}
  \item[$id(T)$] constructs an identity cast from a type
  \item[$seq(c_1,c_2)$] composes two casts
  \item[$cast(T_1,l,T_2)$] constructs a cast from $T_1$ to $T_2$
  \item[$applyCast(v,c)$] applies a cast to a value, producing a cast result
  \end{description}
  We denote by $c : T_1 \Longrightarrow T_2$ to mean $c$ is in $Cast \; T_1 \; T_2$
  and we say ``$c$ is from $T_1$ to $T_2$''.
\end{definition}

@@

Let $v$ range over values. In \ineffCEK{} we have one value
constructor for all castes values.  In \effCEK{C}, however, we do not
have this generic value constructor, instead we push those casts into
the ordinary values. For instance, $(\vOOcons{v_1}{v_2})$ corresponds
to $(\hcvOOcons{v_1'}{id(T_1)}{v_2'}{id(T_2)})$. To construct values
of the dynamic type, we use $\hcvOOinj{P}{v}$.

Cast results ($r$) and machine states ($s$) are the same as for
\ineffCEK{}.

@@

In this paper we presented a new cast representation, hypercoercions, in two 
flavors, Lazy D and Lazy UD. Hypercoercions have a structurally recursive 
composition, which is a win over coercions in normal form for mechanized 
metatheory. What is more, hypercoercions have a more compact memory 
representation. We prove that $S(H)$, an abstract machine that uses Lazy D 
hypercoercions, is equivalent to $D$, a semantics of Lazy D Cast Calculus.

We also present a framework for proving the correctness of Lazy D Cast ADT. 
If $C$ implements the Lazy D Cast ADT, $S(C)$ would be equivalent to $D$.

\todo[inline]{related works. -KC}

>>>>

\newcommand{\judgeCreduce}[2]{#1 \longmapsto_{\mathcal{X}} #2}
\newcommand{\judgeCreduceTrans}[2]{#1 \longmapsto_{\mathcal{X}}^{*} #2}
\newcommand{\judgeCeval}[2]{eval_{\mathcal{X}}(#1) = #2}
\newcommand{\redrule}[3]{#1 & \longmapsto_\mathcal{X} & #2 & #3\\}
\newcommand{\hiredrule}[3]{\highlight{#1} & \highlight{\longmapsto_\mathcal{X}} 

@@

\newcommand{\ineffCEK}{$\mathcal{X}$}

@@

The framework includes an abstract data type named Cast ADT 
(Section~\ref{sec:framework:adt}), and a CEK machine \effCEK{C} that is 
parameterized over Cast ADTs (Section~\ref{sec:framework:cek}). This machine is 
space-efficient provided that the Cast ADT implementation performs compression. 
We conjecture that all cast representations defined in the literature
are instances of this ADT.

%We start by defining a CEK machine \effCEK{C} that is parameterized
%over the cast representation $C$
%(Section~\ref{sec:framework:cek}). This machine is space-efficient
%provided that the cast representation performs compression.  The
%interface between the machine and the cast representation is defined
%by an abstract data type, named Cast ADT, that we define shortly 
%(Section~\ref{sec:framework:adt}).  We
%conjecture that all cast representations defined in the literature
%are instances of this ADT.

@@

Let $v$ range over values. In \ineffCEK{} we have one value
constructor for all casted values.  In \effCEK{C}, however, we do not
have this generic value constructor, instead we push casts into the ordinary 
values. Thus non-casted values in \ineffCEK\ corresponds to values in 
\effCEK{C} where casts are identities. For instance, $(\vOOcons{v_1}{v_2})$ 
corresponds to $(\hcvOOcons{v_1'}{id(T_1)}{v_2'}{id(T_2)})$.

%\hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2} \\
%c_1' = seq(c_1,cast(T_1,l,T_3)) \\
%c_2' = seq(c_2,cast(T_2,l,T_4))
%}{
%\hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
%\vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
%: \POOprod{T_3}{T_4}
%}

Cast results ($r$) are the same as for \ineffCEK.

@@

\begin{definition}[Cast Abstract Data Type (Cast ADT)]
  \label{def:cast-rep}
  A cast abstract data type is a set $Cast$, which is indexed by two types, 
  with four operators:
  \begin{description}
  \item[$id(T)=c$] constructs an identity cast from a type
  \item[$seq(c,c)=c$] composes two casts
  \item[$cast(T,l,T)=c$] constructs a cast from a source type, a label, and a 
  target type
  \item[$applyCast(v,c)=r$] applies a cast to a value, producing a cast result
  \end{description}
  We denote by $c : T_1 \Longrightarrow T_2$ to mean $c$ is in $Cast \; T_1 \; T_2$
  and we say ``$c$ is from $T_1$ to $T_2$''.
\end{definition}

@@

@@

In this paper we presented a new cast representation, hypercoercions, in two 
flavors, Lazy D and Lazy UD. Hypercoercions have a structurally recursive 
composition, which is a win over coercions in normal form for mechanized 
metatheory. What is more, hypercoercions have a more compact memory 
representation.

We also present steps toward the first framework for proving the correctness of 
cast representations.
%
The framework includes abstract data types for cast representations (Cast ADT 
and Lazy D Cast ADT), an parameterized abstract machine \effCEK{C}, and a 
theorem states that \effCEK{C} is equivalent to $D$ if $C$ is a Lazy D Cast 
ADT.
Cast ADT is the interface between cast representations and \effCEK{C}. Its
implementation should provide four functions: $cast(T,l,T) = c$, $id(T)=c$, and 
$seq(c,c)=c$ for cast construction, and $applyCast(v,c)=r$ for cast application.
%
Lazy D Cast ADT refines Cast ADT. We conjecture that Lazy D threesomes, and 
Lazy D coercions in normal form are also Lazy D Cast ADT.

Finally, we proved that Lazy D hypercoercions ($H$) is a Lazy D Cast ADT. By 
using our framework, we shown that it is a correct cast representation because 
$S(H)$ is equivalent to $D$.
